---
title: "Sensitivity_Analysis_3"
author: "Yineng Chen"
date: "2020/6/2"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lme4)
library(sjPlot)
library(arsenal)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```



Data from 11 patients who have treatment intervention were removed. 
Data from the remaining 23 patients including those who were younger than 18 at baseline were analyzed using mixed-effect models which assume that weight change is linear over time.


```{r,include=FALSE}
#########################################
# load data
weight_data = read_csv("./data/dataclean2.csv",col_types = "ccfdddcdffff") %>% 
  janitor::clean_names() %>% drop_na(id)
# View(weight_data)
# dim(weight_data)

### process data ####
### data befroe intervention ###
weight_before = weight_data %>% 
  filter(tnp == 0 & feeding_tube == 0 & bm_transplant == 0 & l_transplant == 0) %>% 
  mutate(years = age - first_age) 

weight_before2 = weight_before %>% 
  select(-dob,-age,-dom) %>% 
  select(id,num_measure,years,weight,everything()) %>% 
  mutate(years = as.numeric(years),weight = as.numeric(weight)) # id = as character

ids = unique(weight_before2$id)
# length(ids) # 24 subjects
### subjuct only have one data before intervention ###
no_id = data.frame(count(weight_before2, id))  # number of data point before intervention
no_1 = filter(no_id, n == 1) # 98 only has one data point
weight_before2 = weight_before2 %>% filter(!(id == 98))  # No.98 is excluded, 23 subjects left
# View(weight_before2)
ids = unique(weight_before2$id) #23 subjects

### seperate by id ###
# Group 1
weight_before2_c1 = weight_before2 %>% 
  mutate(id = as.numeric(id)) %>% 
  filter(id < 200)
# View(weight_before2_c1)
# unique(weight_before2_c1$id)
# Group 2
weight_before2_c2 = weight_before2 %>%
  mutate(id = as.numeric(id)) %>% 
  filter(id > 200)
# View(weight_before2_c2)
# unique(weight_before2_c2$id)
```


## Plots

```{r,echo = FALSE}
#####################################################
# speghetti plot

Weight_plot = ggplot(data = weight_before2, aes(x = years, y = weight, group = id)) +
  geom_line(aes(color = id)) +
  geom_point(aes(color = id)) +
  scale_color_hue(name = "Subjects", h = c(100, 300)) +
  ggtitle("Weight changes over time (years) for patients in cohort 1 and 2")+ 
  labs(x = "Time (years)", y = "Weight (kg)")+
  theme_bw() + 
  theme(legend.position = "bottom")

# individually

# for c1
#dev.off()
par(mar = c(4,4,2,2))
par(mfrow = c(3,3))
idc1 = unique(weight_before2_c1$id)
for(i in idc1){
  df = subset(weight_before2_c1, weight_before2_c1$id == i)
  plot( df$years, df$weight,main = i , type = "b",
        xlab = "Time (years)",ylab = "weight (kg)")
}


idc2 = unique(weight_before2_c2$id)
#par(mar = c(4,4,2,2))
#par(mfrow = c(3,3))
for(i in idc2){
  df = subset(weight_before2_c2, weight_before2_c2$id == i)
  plot( df$years, df$weight, main = i, 
        type = "b",xlab = "Time (years)",ylab = "weight (kg)")
}

Weight_plot
```


# Baseline

```{r,echo = FALSE}
baseline = read_csv("./data/baseline2.csv",col_types = "cfdd") %>% 
  janitor::clean_names() %>% drop_na(id)

baseline <- baseline %>% 
  mutate(cohort=factor(cohort, labels=c("cohort 1","cohort 2")))
# First table 
my_labels <- list(age = "Age (years)", weight = "Weight(kg)", cohort ="Cohort")

my_controls <- tableby.control(
  total = T,
  test= T,  # No test p-values yet
  numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),
  cat.stats = c("countpct", "Nmiss2"),
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max",
    Nmiss2 = "Missing",
    countpct = "N (%)"))

tab1 <- tableby(cohort ~ age + weight , data=baseline, control=my_controls)  

summary(tab1,title = "Descriptive Statistics at baseline", labelTranslations = my_labels, text=T)

```


# Mixed effect models

## cohort 1 and 2 

The weight loss rate in all 23 MNGIE patients is 0.82 kg/year (n=23, SE=0.09). 


```{r,echo = FALSE}

#############################
#     Mixed effect models   #
#############################

# FIt a model fot cohort 1 and 2
mod = lme4::lmer(weight ~ years+ (1|id), REML = T, data=weight_before2)
# summary(mod)
# sjPlot::plot_model(mod,
#                   show.values = TRUE, show.p=TRUE,
#                   title = "Weight changes over time (years)")
sjPlot:: tab_model(mod,show.re.var= F, 
                  dv.labels= "Weight changes over time (years) in cohort 1 and 2")
effects_years <- effects::effect(term= "years", mod = mod, xlevels = list(years = c(0,5,10,15,20,25,27)))
# summary(effects_years) 
# Save the effects values as a df:
x_years <- as.data.frame(effects_years)

years_plot = ggplot() + 
  #2
  geom_line(data=weight_before2, aes(x = years, y = weight,group = id,color = id)) + 
  geom_point(data=weight_before2, aes(x = years, y = weight,group = id,color = id)) + 
  #3
  geom_point(data=x_years, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Weight changes over time (years) in cohort 1 and 2", y="Weight (kg)")

years_plot


```

## cohort 1

The weight loss rate in non-natural history study patients is 0.85 kg/year  (n=14, SE= 0.08).


```{r,echo = FALSE}
##########
# fit a model for cohort 1
mod_c1 = lme4::lmer(weight ~ years+ (1|id),REML= T, data=weight_before2_c1)
# summary(mod_c1)
# sjPlot::plot_model(mod_c1,
#                   show.values = TRUE, show.p=TRUE,
#                   title ="Weight changes over time (years) in cohort 1")
sjPlot:: tab_model(mod_c1,show.re.var= F, 
                   dv.labels= "Weight changes over time (years) in cohort 1")
effects_years_c1 <- effects::effect(term= "years", mod= mod_c1,xlevels = list(years = c(0,5,10,15,20,25,27)))
# summary(effects_years_c1) 
# Save the effects values as a df:
x_years_c1 <- as.data.frame(effects_years_c1)

years_plot_c1 = ggplot() + 
  #2
  geom_line(data=weight_before2_c1 %>% mutate(id = as.character(id)), aes(x = years, y = weight,group = id,color = id)) + 
  geom_point(data=weight_before2_c1 %>% mutate(id = as.character(id)), aes(x = years, y = weight,group = id,color = id)) + 
  #3
  geom_point(data=x_years_c1, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_c1, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_c1, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Weight changes over time (years) in cohort 1", y="Weight (kg)")

years_plot_c1
```


## cohort 2

The weight loss rate in natural history study MNGIE patients is 0.69 kg/year (n=9, SE= 0.22).


```{r,echo = FALSE}
################
# fit a model for cohort 2
mod_c2 = lme4::lmer(weight ~ years+ (1|id),REML= T, data=weight_before2_c2)
# summary(mod_c2)
# sjPlot::plot_model(mod_c2,
#                    show.values = TRUE, show.p=TRUE,
#                    title ="Weight changes over time (years)")
sjPlot:: tab_model(mod_c2,show.re.var= F, 
                   dv.labels= "Weight changes over time (years) in cohort 2")
effects_years_c2 <- effects::effect(term= "years", mod= mod_c2,xlevels = list(years = c(0,5,10,15,19.55)))
# summary(effects_years_c2) 
# Save the effects values as a df:
x_years_c2 <- as.data.frame(effects_years_c2)

years_plot_c2   = ggplot() + 
  #2
  geom_line(data=weight_before2_c2 %>% mutate(id = as.character(id)), aes(x = years, y = weight,group = id,color = id)) + 
  geom_point(data=weight_before2_c2 %>% mutate(id = as.character(id)), aes(x = years, y = weight,group = id,color = id)) +
  #3
  geom_point(data=x_years_c2, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_c2, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_c2, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x ="Weight changes over time (years) in cohort 2", y ="Weight (kg)")+
  theme_bw() + 
  theme(legend.position = "bottom")

years_plot_c2 
```

# effect of cohort

The difference in weight loss between the natural history and non-natural history patients is not significant (p=0.69).


```{r,echo = FALSE}
# A model includs "group"

mod2 = lme4::lmer(weight ~ years+ group+ (1|id),REML= FALSE, data=weight_before2 )
# summary(mod2)
sjPlot:: tab_model(mod2,show.re.var= FALSE, 
                   dv.labels= "Weight change over time (years) in cohort 1 and 2")


```


# truncated models for all patients

```{r, include=False}
# truncated data

weight_chunk = weight_before2 %>% filter(years <= 20)
# view(weight_chunk)
weight_chunk_c1 = weight_chunk %>% 
  mutate(id = as.numeric(id)) %>% 
  filter(id < 200)
weight_chunk_c2 = weight_chunk %>% 
  mutate(id = as.numeric(id)) %>% 
  filter(id > 200) %>% 
  filter(years <= 11)
```
