---
title: 'Sensitivity Analyses'
author: "Yineng Chen"
date: "2020/5/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(nlme)
library(ggplot2)

library(lme4)
library(sjPlot)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

# data from Weight3.R
```



```{r,include=FALSE}
# load data
weight_data = read_csv("./data/dataclean.csv",col_types = "ccfdddcdffff") %>% 
  janitor::clean_names() %>% drop_na(id)
# View(weight_data)
# dim(weight_data)

### process data ####
### data befroe intervention ###
weight_before = weight_data %>% 
  filter(tnp == 0 & feeding_tube == 0 & bm_transplant == 0 & l_transplant == 0) %>% 
  mutate(years = age - first_age) 

weight_before2 = weight_before %>% 
  select(-dob,-age,-first_age,-dom) %>% 
  select(id,num_measure,years,weight,everything()) %>% 
  mutate(years = as.numeric(years),weight = as.numeric(weight)) # id = as character
# View(weight_before2)
ids = unique(weight_before2$id)
# length(ids) # 24 subjects
### subjuct only have one data before intervention ###
no_id = data.frame(count(weight_before2, id))  # number of data point before intervention
no_1 = filter(no_id, n == 1) # 98 only has one data point
weight_before2 = weight_before2 %>% filter(!(id == 98))  # No.98 is excluded, 23 subjects left

ids = unique(weight_before2$id) #23 subjects

### seperate by id ###
# Group 1
weight_before2_c1 = weight_before2 %>% 
  mutate(id = as.numeric(id)) %>% 
  filter(id < 200)
# View(weight_before2_c1)
# unique(weight_before2_c1$id)

# Group 2
weight_before2_c2 = weight_before2 %>%
  mutate(id = as.numeric(id)) %>% 
  filter(id > 200)
# View(weight_before2_c2)
# unique(weight_before2_c2$id)

# chunk data

weight_chunk = weight_before2 %>% filter(years <= 20)
# view(weight_chunk)
weight_chunk_c1 = weight_chunk %>% 
  mutate(id = as.numeric(id)) %>% 
  filter(id < 200)
weight_chunk_c2 = weight_chunk %>% 
  mutate(id = as.numeric(id)) %>% 
  filter(id > 200) %>% 
  filter(years <= 11)
```

## cohort 1 and 2

In this model data from both cohort 1 and 2 were used. Data collected after the 20th year of follow up were excluded.

### Sensitivity Analysis Results

```{r,echo = FALSE}
# cohrt 1 and 2
mod_sen = lme4::lmer(weight ~ years+ (1|id),REML= FALSE, data=weight_chunk)
# summary(mod_sen)

sjPlot:: tab_model(mod_sen,show.re.var= F, 
                   dv.labels= "Sensitivity analysis: Weight changes over time (years) in cohort 1 and 2")
effects_years_sen <- effects::effect(term= "years", mod= mod_sen)
# summary(effects_years_sen) 
# Save the effects values as a df:
x_years_sen <- as.data.frame(effects_years_sen)

years_plot_sen = ggplot() + 
  #2
  geom_line(data=weight_chunk, aes(x = years, y = weight,group = id)) + 
  #3
  geom_point(data=x_years_sen, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_sen, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_sen, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Sensitivity analysis: Weight changes over time (years) in cohort 1 and 2", y="Weight (kg)")


years_plot_sen

```

### Main Analysis Results

```{r,echo = FALSE}
mod = lme4::lmer(weight ~ years+ (1|id),REML= FALSE, data=weight_before2)
sjPlot:: tab_model(mod,show.re.var= F, 
                  dv.labels= "Weight changes over time (years) in cohort 1 and 2")

```



## cohort 1

In this model data from cohort 1 were used. Data collected after the 20th year of follow up were excluded.

### Sensitivity Analysis Results

```{r,echo = FALSE}
mod_sen_c1 = lme4::lmer(weight ~ years+ (1|id),REML= FALSE, data=weight_chunk_c1)
# summary(mod_sen_c1)

sjPlot:: tab_model(mod_sen_c1,show.re.var= F, 
                  dv.labels= "Sensitivity analysis: Weight changes over time (years) in cohort 1")
effects_years_sen_c1 <- effects::effect(term= "years", mod= mod_sen_c1)
# summary(effects_years_sen_c1) 
# Save the effects values as a df:
x_years_sen_c1 <- as.data.frame(effects_years_sen_c1)

years_plot_sen_c1 = ggplot() + 
  #2
  geom_line(data=weight_chunk_c1, aes(x = years, y = weight,group = id)) + 
  #3
  geom_point(data=x_years_sen_c1, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_sen_c1, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_sen_c1, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Sensitivity analysis: Weight changes over time (years) in cohort 1", y="Weight (kg)")

years_plot_sen_c1
```

### Main Analysis Results

```{r,echo = FALSE}
mod_c1 = lme4::lmer(weight ~ years+ (1|id),REML= FALSE, data=weight_before2_c1)
sjPlot:: tab_model(mod_c1,show.re.var= F, 
                   dv.labels= "Weight changes over time (years) in cohort 1")
```



## cohort 2

In this model data from cohort 2 were used. Data collected after the 11th year of follow up were excluded.

```{r,echo = FALSE}
# cohort 2
mod_sen_c2 = lme4::lmer(weight ~ years+ (1|id),REML= FALSE, data=weight_chunk_c2)
# summary(mod_sen_c2)

sjPlot:: tab_model(mod_sen_c2,show.re.var= F, 
                   dv.labels= "Sensitivity analysis: Weight changes over time (years) in cohort 2")
effects_years_sen_c2 <- effects::effect(term= "years", mod= mod_sen_c2)
# summary(effects_years_sen_c2) 
# Save the effects values as a df:
x_years_sen_c2 <- as.data.frame(effects_years_sen_c2)

years_plot_sen_c2 = ggplot() + 
  #2
  geom_line(data=weight_chunk_c2, aes(x = years, y = weight,group = id)) + 
  #3
  geom_point(data=x_years_sen_c2, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_sen_c2, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_sen_c2, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Sensitivity analysis: Weight changes over time (years) in cohort 2", y="Weight (kg)")


years_plot_sen_c2 
```

### Main Analysis Results

```{r,echo = FALSE}
mod_c2 = lme4::lmer(weight ~ years+ (1|id),REML= FALSE, data=weight_before2_c2)
sjPlot:: tab_model(mod_c2,show.re.var= F, 
                   dv.labels= "Weight changes over time (years) in cohort 2")
```

## Model includs group

```{r,echo = FALSE}
# A model includs "group"

mod2 = lme4::lmer(weight ~ years+ group+ (1|id),REML= FALSE, data=weight_before2 )
# summary(mod2)
sjPlot:: tab_model(mod2,show.re.var= FALSE, 
                   dv.labels= "Weight change over time (years) in cohort 1 and 2")
```

