---
title: "Weight"
author: "Yineng Chen"
date: "2020/5/20"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(nlme)
library(ggplot2)

library(cowplot) 
library(lme4)
library(sjPlot)
library(sjmisc) 
library(effects)
library(sjstats)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```



```{r, include=FALSE}
# load data
weight_data = read_csv("./data/dataclean.csv",col_types = "ccdddcdffff") %>% 
  janitor::clean_names() %>% drop_na(id)

#############################
years = weight_data$age - weight_data$first_age 

weight_data1 = select(weight_data, -dob,-age,-first_age,-dom) %>% 
  mutate(years = years) %>% select(id,num_measure,years,weight,everything()) %>% 
  mutate(id = as.numeric(id),years = as.numeric(years),weight = as.numeric(weight))
# View(weight_data1)

# data befroe intervention
weight_before = weight_data1 %>% filter(tnp == 0 & feeding_tube == 0 & bm_transplant == 0 & l_transplant == 0)
ids = unique(weight_before$id)
length(ids) # 24 subjects

# subjuct only have one data before intervention
no_id = data.frame(count(weight_before, id))  # number of data point before intervention
no_1 = filter(no_id, n == 1) # 98 only has one data point
weight_before2 = weight_before %>% filter(!(id == 98))  # 23 subjects left

#seperate by id
weight_before2_c1 = weight_before2 %>% filter(id < 200)
# View(weight_before2_c1)
# unique(weight_before2_c1$id)

weight_before2_c2 = weight_before2 %>% filter(id > 200)
# View(weight_before2_c2)
# unique(weight_before2_c2$id)

```

There are 32 subjects with valid data in total. Bodyweight after any intervention (TPN, Feeding tube, transplant) is not included in this data set. 23 subjects have more than one valid data point before any intervention. Cohort 1 contains patients who are not in the natural history study with id smaller than 200. Cohort 2 includes patients coming from natural history study with id larger than 200.

Note: 

1. patien No.206  2 years after the end of TPN use

1. patien No.33 3 years after the end of  potential TPN use

Body weight changes for each subject over time are shown below.

years = date of the least measurement - date of the first measurement


```{r,echo = FALSE}
# plot individually
par(mar = c(4,4,2,2))
par(mfrow = c(3,3))
idc1_1 = unique(weight_before2_c1$id)[1:9]
for(i in idc1_1){
  df = subset(weight_before2_c1, weight_before2_c1$id == i)
  plot( df$years, df$weight,main = i , type = "b")
}

par(mfrow = c(3,3))
idc1_2 = unique(weight_before2_c1$id)[10:14]
for(i in idc1_2){
  df = subset(weight_before2_c1, weight_before2_c1$id == i)
  plot( df$years, df$weight,main = i , type = "b")
}

par(mar = c(4,4,2,2))
par(mfrow = c(3,3))
idc2 = unique(weight_before2_c2$id)

for(i in idc2){
  df = subset(weight_before2_c2, weight_before2_c2$id == i)
  plot( df$years, df$weight, main = i, 
        type = "b",xlab = "years",ylab = "weight(kg)")
}

# speghetti plot
data_ggp = weight_before2 %>%  mutate(id = as.character(id))
Weight_plot = ggplot(data = data_ggp, aes(x = years, y = weight, group = id)) +
  geom_line(aes(color = id)) +
  scale_color_hue(name = "Sunjects", h = c(100, 300)) +
  ggtitle("Weight change over time(year)")
  

Weight_plot
```





```{r,include=FALSE}
# fit LMM with random intercept 
#LMM1 <- lme(weight ~ years, random = ~  1 | id, data = weight_before2)
#summary(LMM1)
#random.effects(LMM1)
#fixed.effects(LMM1)
#weight_before2$fit= fitted(LMM1)
######################

#################

```

##regrression Model

Fit a regression model for all patients.the mean of bodyweight decrease 0.81 kg per year (SD = 0.08381, 95% CI(-0.97 ~ -0.64)).

```{r, echo = FALSE}
# fit LMM with random intercept for all data
mod = lme4::lmer(weight ~ years + (1|id),REML= FALSE, data=weight_before2 )
# summary(mod)
# sjPlot::plot_model(mod,
#                   show.values = TRUE, show.p=TRUE,
#                   title ="Weight change over time(years)")
sjPlot:: tab_model(mod,show.re.var= TRUE, 
                   dv.labels= "Weight change over time(years)")
effects_years <- effects::effect(term= "years", mod= mod)
# summary(effects_years)

# Save the effects values as a df:
x_years <- as.data.frame(effects_years)

years_plot = ggplot() + 
  #2
  geom_line(data=data_ggp, aes(x = years, y = weight,group = id)) + 
  #3
  geom_point(data=x_years, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="time(years)", y="Weight(kg)") +
  ggtitle("Weight change over time(year)")

years_plot
```

###cohort 1 

Fit a regression model for patients from cohort 1. In general, the body weight decrease 0.85 kg per year (SE = 0.08101, 95% CI(-1.01 ~ -0.69)).

```{r,echo = FALSE}
# fit LMM with random intercept for all data
mod_c1 = lme4::lmer(weight ~ years + (1|id),REML= FALSE, data=weight_before2_c1)
# summary(mod_c1)
# sjPlot::plot_model(mod,
#                   show.values = TRUE, show.p=TRUE,
#                   title ="Weight change over time(years)")
sjPlot:: tab_model(mod_c1,show.re.var= TRUE, 
                   dv.labels= "Weight change over in chort 1")
effects_years_c1 <- effects::effect(term= "years", mod= mod_c1)
# summary(effects_years)

# Save the effects values as a df:
x_years_c1 <- as.data.frame(effects_years_c1)

years_plot_c1 = ggplot() + 
  #2
  geom_line(data=weight_before2_c1, aes(x = years, y = weight,group = as.character(id))) + 
  #3
  geom_point(data=x_years_c1, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_c1, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_c1, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="time(years)", y="Weight(kg)") +
  ggtitle("Weight change in cohort 1")


```

###chort 2(id > 200)

Fit a regression model for patients from cohort 2. In general, the body weight decrease 0.66 kg per year (SE = 0.1974, 95% CI(-1.04 ~ -0.27)).

```{r, echo = FALSE}
mod_c2 = lme4::lmer(weight ~ years + (1|id),REML= FALSE, data=weight_before2_c2)
# summary(mod_c2)
# sjPlot::plot_model(mod,
#                   show.values = TRUE, show.p=TRUE,
#                   title ="Weight change over time(years)")
sjPlot:: tab_model(mod_c2,show.re.var= TRUE, 
                   dv.labels= "Weight change over time(years) in chort 2")
effects_years_c2 <- effects::effect(term= "years", mod= mod_c2)
# summary(effects_years)

# Save the effects values as a df:
x_years_c2 <- as.data.frame(effects_years_c2)

years_plot_c2 = ggplot() + 
  #2
  geom_line(data=weight_before2_c2, aes(x = years, y = weight,group = as.character(id))) + 
  #3
  geom_point(data=x_years_c2, aes(x=years, y=fit), color="blue") +
  #4
  geom_line(data=x_years_c2, aes(x=years, y=fit), color="blue") +
  #5
  geom_ribbon(data= x_years_c2, aes(x=years, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Wtime(years)", y="Weight(kg)") +
  ggtitle("Weight change over time(year) in cohort 2")


```

```{r,echo = FALSE}
years_plot_c1 + years_plot_c2 
```

